%
% Chapter 3 — System Requirements
%
\chapter{System Requirements} \label{cap:requirements}

This chapter presents the requirements of the \textit{Diário LX} platform, derived from the
project specification and validated with the editorial team. Section~\ref{sec:req_functional}
sets out the functional requirements --- from the content lifecycle and taxonomy to user
permissions, the two site faces, and search; and Section~\ref{sec:req_nonfunctional} covers
the non-functional ones, such as security, media handling, and responsiveness.

\section{Functional Requirements} \label{sec:req_functional}

\subsection{Content Management} \label{sec:req_content}

The system must provide a full content management lifecycle supporting the following content types: \textbf{articles}, \textbf{podcasts}, \textbf{podcast episodes}, \textbf{videos}, and \textbf{photo essays}. Each content item must include the following attributes:

\begin{itemize}
      \item Title
      \item Body text (supporting inline subtitles, highlights, images, and captioned image galleries)
      \item Lead paragraph (\textit{entrada})
      \item Author(s) and photo author
      \item Creation and publication date
      \item Featured image
      \item Category and subcategory
      \item Tags
\end{itemize}

Content must progress through a defined workflow:

\begin{enumerate}
      \item \textbf{Draft} — created and edited by a Contributor.
      \item \textbf{Pending Review} — submitted for editorial review.
      \item \textbf{Approved} — endorsed by an Editor or Administrator and published to
            readers, either immediately or at a scheduled time.
      \item \textbf{Rejected} — returned to the author with revision comments.
\end{enumerate}

It must be possible to archive and permanently delete content. It must be possible to mark content as featured for display on the homepage.

\subsection{Categories, Subcategories, and Tags} \label{sec:req_categories}

Content must be organised using categories and subcategories. The main editorial sections are:\footnote{Section names are kept in
      Portuguese as they are proper editorial names specific to \textit{Diário LX}.}
\begin{itemize}
      \item \textit{Lisboa, Cidade Aberta}
      \item \textit{A Fundo}
      \item \textit{Secções} (\textit{Mundo, Política, Sociedade, Economia, Cultura e Media, Bem-Viver, Desporto})
      \item \textit{Especiais}
      \item \textit{Fotografia}
      \item \textit{Podcasts}
      \item \textit{Vídeos}
\end{itemize}

Each category must support a name, an associated colour, and an automatically generated listing page. Tags must also have dedicated listing pages. The system must allow the creation, editing, and deletion of categories and tags.

\subsection{User Management and Permissions} \label{sec:req_users}

The system defines three user roles:

\begin{description}
      \item[Administrator] The highest privilege level. Inherits all Editor and Contributor
            permissions. Additionally manages user accounts and generates invitation codes for
            new users to register on the platform.
      \item[Editor] Inherits all Contributor permissions. Can approve or reject content
            submitted for review.
      \item[Contributor] The base role. Can create content and submit it for editorial
            review. Cannot publish directly.
\end{description}

The three roles form a strict permission hierarchy:
\textbf{Contributor} $\subset$ \textbf{Editor} $\subset$ \textbf{Administrator},
where each role inherits all permissions of the roles below it.

Each user profile includes: name, email, profile photograph, active/inactive status, a position, a biography, and role.

\subsection{Backoffice} \label{sec:req_backoffice}

The backoffice must provide:

\begin{itemize}
      \item Account creation for Editors, Contributors and Administrators via invitation codes.
      \item Login and logout for all profiles.
      \item Listing and search of all content and users.
      \item Profile editing (own profile for all; any profile for Administrators).
      \item Category and tag management (create, edit, delete) by Administrators.
\end{itemize}

\subsection{Public Website} \label{sec:req_frontend}

\subsubsection{Homepage}
The homepage must display:
\begin{itemize}
      \item Four featured articles (defined in the backoffice).
      \item The four most recent articles.
      \item Navigation to main editorial sections.
      \item Links to institutional pages: Team, Editorial Statute, and Code of Ethics.
\end{itemize}

\subsubsection{Other Pages}
\begin{itemize}
      \item \textbf{Category page}: articles in reverse chronological order.
      \item \textbf{Tag page}: list of associated articles.
      \item \textbf{Article page}: full article view.
      \item \textbf{Team page}.
      \item \textbf{Podcast page}: dedicated layout distinct from other category pages.
\end{itemize}

\subsection{Search and Navigation} \label{sec:req_search}

\begin{itemize}
      \item Keyword search across articles.
      \item Human-readable, slugified URLs (e.g. \texttt{/p/orcamento-estado-2026}).
\end{itemize}

\section{Non-Functional Requirements} \label{sec:req_nonfunctional}

\begin{itemize}
      \item \textbf{Security}: Authentication via JSON Web Tokens (JWT), with short-lived
            access tokens and long-lived refresh tokens both delivered as \texttt{HttpOnly}
            cookies. Role-based access control enforced at the API level.

      \item \textbf{Media handling}: The platform must accept media uploads of up to two
            gigabytes per file, covering the high-resolution imagery, audio, and video used
            in editorial content.

      \item \textbf{Maintainability}: The system is organised into independent modules
            (HTTP, Services, Domain, Repository, and Storage) allowing the frontend and backend
            to evolve independently.

      \item \textbf{Compatibility}: The public website must be fully functional on mobile,
            tablet, and desktop viewports. The backoffice is optimised for desktop use.
\end{itemize}
