%
% Chapter 5 — Data Model
%
\chapter{Data Model} \label{cap:datamodel}

\section{Entities} \label{sec:dm_entities}

This section describes the core entities of the \textit{Diário LX} data model, their
attributes, and the relationships between them. The complete entity-relationship
diagram is presented in Appendix~\ref{app:db_model}.

\subsection{Enumerated Types} \label{sec:dm_enums}

Several columns are typed using PostgreSQL enumerated types, which constrain the set
of admissible values at the database level:

\begin{description}
      \item[\texttt{user\_role}] \textit{ADMIN}, \textit{EDITOR}, \textit{CONTRIBUTOR}.

      \item[\texttt{content\_type}] \textit{ARTICLE}, \textit{VIDEO}, \textit{EPISODE},
            \textit{PODCAST}, \textit{PHOTO\_ESSAY}.

      \item[\texttt{content\_state}] \textit{DRAFT}, \textit{PENDING\_REVIEW},
            \textit{PUBLISHED}, \textit{REJECTED} --- the editorial workflow states of a
            content item.

      \item[\texttt{content\_history\_action}] \textit{APPROVED}, \textit{REJECTED} ---
            the moderation actions recorded in the review history.

      \item[\texttt{block\_type}] \textit{TEXT}, \textit{QUOTE}, \textit{MEDIA},
            \textit{H3}, \textit{H4}, \textit{GALLERY}, \textit{EMBED}.

      \item[\texttt{media\_purpose}] \textit{GALLERY} (editorial media) or
            \textit{PROFILE} (user avatars).

      \item[\texttt{credit\_role}] one of \textit{PHOTOGRAPHER}, \textit{VIDEOGRAPHER},
            \textit{AUDIO\_ENGINEER}, \textit{WRITER}, \textit{DIRECTOR}, \textit{PRODUCER},
            \textit{EDITOR}, \textit{NARRATOR}, \textit{ACTOR}, \textit{MODEL},
            \textit{ILLUSTRATOR}, \textit{DESIGNER}, or \textit{OTHER}.
\end{description}

\subsection{User} \label{sec:dm_user}

Represents an authenticated system user. Each user has a \texttt{role}
(\textit{ADMIN}, \textit{EDITOR}, or \textit{CONTRIBUTOR}), credentials
(\texttt{username}, \texttt{email}, \texttt{password\_hash}), and profile information
(\texttt{first\_name}, \texttt{last\_name}, \texttt{position}, \texttt{bio}), together
with an optional \texttt{avatar\_media\_id} reference to a media item. An
\texttt{on\_team} flag marks users that appear on the public team page, and an
\texttt{active\_account} flag is used to deactivate access without deleting the record.
Creation and modification are tracked through \texttt{created\_at} and
\texttt{updated\_at}.

\subsection{Session} \label{sec:dm_session}

Stores active refresh tokens issued upon login. Each session is associated with a
user and holds a \texttt{refresh\_token}, \texttt{created\_at}, and
\texttt{expires\_at}. Sessions are deleted on logout and expire automatically based
on \texttt{expires\_at}.

\subsection{Invite} \label{sec:dm_invite}

Represents an invitation token generated by an Administrator to allow new user
registration. Each invite carries an \texttt{invite\_token}, a \texttt{role\_assigned},
an expiration timestamp, and a \texttt{used} flag to prevent reuse.

\subsection{Content} \label{sec:dm_content}

The central entity of the system. Each content item has a \texttt{type}
(\textit{ARTICLE}, \textit{VIDEO}, \textit{EPISODE}, \textit{PODCAST}, or
\textit{PHOTO\_ESSAY}), a \texttt{title}, a \texttt{headline}, a \texttt{slug}, and an
optional reference to its \texttt{category}. The body of a content item is composed of
an ordered sequence of \texttt{content\_blocks}.

The editorial lifecycle of a content item is tracked by the \texttt{state} column
(\texttt{content\_state}), which is the primary control over public visibility. Content
progresses through \textit{DRAFT}, \textit{PENDING\_REVIEW}, and \textit{PUBLISHED},
or is \textit{REJECTED} during review; only \textit{PUBLISHED} items are exposed to
readers. The \texttt{published\_at} timestamp records the moment of the first
publication and is preserved across subsequent edits and re-publications, so that it
always reflects the original publication date; the \texttt{updated\_at} timestamp, in
turn, captures the most recent modification and is surfaced publicly as an
``updated on'' indicator. A future \texttt{published\_at} value additionally allows a
content item to be withheld from listings until that instant is reached.

Two content types rely on additional columns. An \textit{EPISODE} references its
parent \textit{PODCAST} through the self-referencing \texttt{parent\_id} foreign key
(\texttt{ON DELETE CASCADE}, so deleting a podcast removes its episodes). \textit{VIDEO}
and \textit{EPISODE} items may embed external media (a YouTube video or a Spotify
episode) through the \texttt{embed\_url} column instead of an uploaded file. Content
items may also be soft-archived via an \texttt{archived\_at} timestamp.

Content items may have one or more authors, recorded in the \texttt{content\_authors}
association table with a \texttt{role} of \textit{primary} or \textit{secondary}.
A partial unique index ensures that each content item has at most one primary author.

\subsection{Content Block} \label{sec:dm_block}

Represents a single unit of content within a content item. Each block has a
\texttt{type} (\textit{TEXT}, \textit{QUOTE}, \textit{H3}, \textit{H4},
\textit{EMBED}, \textit{MEDIA}, or \textit{GALLERY}), a \texttt{position} defining its
order within the content item, and, depending on the type, either a \texttt{content}
text field or a \texttt{media\_id} reference. A database constraint enforces this:
textual blocks (\textit{TEXT}, \textit{QUOTE}, \textit{H3}, \textit{H4},
\textit{EMBED}) must have \texttt{content} set, \textit{MEDIA} blocks must reference a
\texttt{media\_id}, and \textit{GALLERY} blocks must have neither (their images are
stored separately). A unique constraint on \texttt{(content\_id, position)} keeps the
ordering consistent.

\subsection{Content Block Images} \label{sec:dm_block_images}

\textit{GALLERY} blocks aggregate an ordered set of images in the
\texttt{content\_block\_images} table. Each row links a block to a \texttt{media\_id},
with an optional \texttt{caption} and a \texttt{position}, keyed by
\texttt{(block\_id, position)}. Rows are removed automatically when the parent block is
deleted.

\subsection{Content History} \label{sec:dm_history}

Records the moderation trail of a content item. Each entry in the
\texttt{content\_history} table references the \texttt{content\_id}, the user who
performed the action (\texttt{performed\_by}, set to \texttt{NULL} if that user is
later removed), the \texttt{action} taken (\textit{APPROVED} or \textit{REJECTED}), an
optional review \texttt{comment}, and the \texttt{performed\_at} timestamp. This
provides an auditable log of the approval workflow and supplies the mandatory comment
attached to rejections.

\subsection{Category} \label{sec:dm_category}

Editorial categories support a two-level hierarchy via a \texttt{parent\_id}
self-reference on the \texttt{categories} table. Top-level categories have
\texttt{parent\_id = NULL}; subcategories reference their parent. Each category
has a \texttt{name}, a unique \texttt{slug}, a unique \texttt{color}, and an
optional \texttt{description}. The foreign key uses \texttt{ON DELETE SET NULL}, so
deleting a parent category promotes its subcategories to top level rather than
removing them. Categories may be soft-deleted via an \texttt{archived\_at} timestamp.

\subsection{Tag} \label{sec:dm_tag}

Tags provide cross-cutting thematic classification. Each tag has a \texttt{name}, a
unique \texttt{slug}, and an optional \texttt{description}, and may be soft-deleted via
an \texttt{archived\_at} timestamp. A content item may have multiple tags, recorded in
the \texttt{content\_tags} association table. Each tag association carries a
\texttt{role} of \textit{primary} or \textit{secondary}; a partial unique index
ensures that each content item has at most one primary tag.

\subsection{Media} \label{sec:dm_media}

Stores metadata for uploaded media files (images, videos, and audio). Binary files
are not stored in the database --- only their \texttt{purpose}
(\textit{GALLERY} or \textit{PROFILE}), \texttt{original\_file\_name}, storage location
(\texttt{bucket} and \texttt{object\_key}, with optional \texttt{thumbnail\_bucket} and
\texttt{thumbnail\_object\_key}), \texttt{alt\_text}, \texttt{mime\_type},
\texttt{size\_bytes}, and a \texttt{status} field (\textit{pending} or \textit{ready})
are persisted. The \textit{pending} state indicates that the frontend has initiated the
upload but not yet confirmed it; the backend transitions the status to \textit{ready}
upon confirmation and records the \texttt{uploaded\_at} timestamp.

Media items may have associated credits via the \texttt{media\_credits} table,
which records the contributing user and their \texttt{credit\_role} (e.g.\
\textit{PHOTOGRAPHER}, \textit{VIDEOGRAPHER}, \textit{AUDIO\_ENGINEER}, \textit{WRITER},
\textit{DIRECTOR}), keyed by \texttt{(media\_id, user\_id, role)} so that a single
person may hold more than one role on the same item.

\subsection{Featured Sections and Contents} \label{sec:dm_featured}

The homepage layout is curated through two related tables. The
\texttt{featured\_sections} table defines the ordered sections of the homepage; each
row has a \texttt{type} (e.g.\ \texttt{HIGHLIGHT}, \texttt{FEATURED}, \texttt{CATEGORY},
\texttt{CATEGORY\_ROW}, \texttt{PHOTOS}, \texttt{PODCASTS}, \texttt{VIDEOS}), a unique
\texttt{position}, and, for \texttt{CATEGORY} sections only, a \texttt{category\_id}
reference. A check constraint enforces that a \texttt{category\_id} is present exactly
when the section type is \texttt{CATEGORY}.

The \texttt{featured\_contents} table then holds the content items placed within each
section: each row references a \texttt{section\_id} and a \texttt{content\_id}, together
with a \texttt{position} for ordering. Unique constraints on
\texttt{(section\_id, content\_id)} and \texttt{(section\_id, position)} prevent
duplicate or conflicting placements, and both foreign keys cascade on deletion.

\subsection{Settings} \label{sec:dm_settings}

Site-wide configuration is stored in a generic key/value \texttt{settings} table,
mapping a \texttt{key} (primary key) to a textual \texttt{value}. It holds
configuration such as social media links, contact information, and the public
navigation layout, allowing these values to be edited by administrators without
schema changes.

\section{Database Views} \label{sec:dm_views}

Several PostgreSQL views are defined to encapsulate complex query logic and provide
pre-joined representations of the data for use by the repository layer:

\begin{description}
      \item[\texttt{v\_contents}] Full content view, including category, featured media,
            authors ordered by role, tags, and content blocks with their associated media
            and gallery images. Used for single content item retrieval.

      \item[\texttt{v\_published\_contents}] Restricts \texttt{v\_contents} to publicly
            visible content (states \textit{PUBLISHED} and \textit{PENDING\_REVIEW}),
            backing the public article pages.

      \item[\texttt{v\_contents\_summary}] Lightweight content listing view, including
            only the fields required for listing pages: title, slug, type, state,
            category, featured image URL, primary tag, and author names. Used for
            paginated content listings in the backoffice.

      \item[\texttt{v\_published\_contents\_summary}] The public counterpart of
            \texttt{v\_contents\_summary}, restricted to visible content for the reader-facing
            listings.

      \item[\texttt{v\_content\_review\_history}] Joins \texttt{content\_history} with the
            performing user to expose the moderation trail (action, comment, author, and
            timestamp) of each content item.

      \item[\texttt{v\_featured\_sections}] Assembles the homepage layout, joining each
            featured section with its category (when applicable) and its ordered list of
            featured content items.

      \item[\texttt{v\_medias}] Media view that computes the full \texttt{url} and
            \texttt{thumbnail\_url} from \texttt{bucket} and \texttt{object\_key}, and
            aggregates credits as a JSON array.

      \item[\texttt{v\_categories}] Category view that joins each category with its
            parent name and includes a \texttt{quantity} count of associated content items.

      \item[\texttt{v\_tags}] Tag view that includes a \texttt{quantity} count of
            associated content items.

      \item[\texttt{v\_users}] User view exposing the public profile fields and joining
            the avatar media so that team and author pages can be served directly.
\end{description}

\section{Indexes} \label{sec:dm_indexes}

The following indexes were defined to support the most frequent query patterns:

\begin{description}
      \item[\texttt{idx\_content\_category\_id}] Supports filtering content by category.

      \item[\texttt{idx\_content\_blocks\_content\_position}] Composite index on
            \texttt{(content\_id, position)}, supporting ordered block retrieval per
            content item.

      \item[\texttt{idx\_content\_authors\_content\_id} and \texttt{idx\_content\_authors\_author\_id}]
            Support join queries on the \texttt{content\_authors} association table.

      \item[\texttt{idx\_content\_tags\_content\_id} and \texttt{idx\_content\_tags\_tag\_id}]
            Support join queries on the \texttt{content\_tags} association table.

      \item[\texttt{idx\_contents\_section}] Composite index on
            \texttt{(section\_id, position)} of \texttt{featured\_contents}, supporting
            ordered retrieval of the items within each homepage section.

      \item[\texttt{uq\_content\_primary\_author}] Partial unique index ensuring that
            each content item has at most one author with \texttt{role = 'primary'}.

      \item[\texttt{uq\_content\_primary\_tag}] Partial unique index ensuring that
            each content item has at most one tag with \texttt{role = 'primary'}.
\end{description}

\section{Key Design Decisions} \label{sec:dm_decisions}

\begin{itemize}
      \item \textbf{Block-based content model}: Instead of storing article body as a
            single rich-text blob, content is decomposed into an ordered sequence of
            typed blocks (text, headings, quotes, embeds, single media, and galleries).
            This enables structured rendering, easier reordering, and per-block media
            references. Gallery images are normalised into a dedicated
            \texttt{content\_block\_images} table.

      \item \textbf{Explicit editorial workflow}: Rather than inferring status from
            timestamps, a dedicated \texttt{content\_state} enumeration models the
            editorial state machine (\textit{DRAFT}, \textit{PENDING\_REVIEW},
            \textit{PUBLISHED}, \textit{REJECTED}). State drives visibility and
            authorisation, while \texttt{published\_at} is preserved as the immutable
            first-publication date and \texttt{updated\_at} tracks later edits. The
            \texttt{content\_history} table provides an auditable record of approvals and
            rejections.

      \item \textbf{Self-referencing hierarchies}: A single \texttt{categories} table
            with a \texttt{parent\_id} foreign key supports the two-level category
            hierarchy, and the same technique on \texttt{contents} links a podcast to its
            episodes without dedicated child tables.

      \item \textbf{Uploaded or embedded media}: Video and episode content may either
            reference an uploaded file or an external provider through \texttt{embed\_url}
            (YouTube or Spotify), avoiding the need to host large third-party media.

      \item \textbf{Two-table homepage curation}: The homepage is modelled as
            \texttt{featured\_sections} (ordered, typed layout slots) and
            \texttt{featured\_contents} (the ordered items within each slot), which keeps
            section configuration independent from the content assigned to it.

      \item \textbf{Media metadata separation}: Binary files are stored in SeaweedFS;
            only metadata is persisted in the database. The \textit{pending}/\textit{ready}
            status tracks the asynchronous upload lifecycle, ensuring that incomplete
            uploads do not surface in the application before the file is available in
            storage.

      \item \textbf{Key/value site settings}: Editable, non-structural configuration
            (social links, contact details, navigation) is stored in a generic
            \texttt{settings} table, allowing administrators to change it at runtime
            without schema migrations.

      \item \textbf{Timestamps as epoch integers}: All timestamps are stored as
            \texttt{BIGINT} Unix epoch values rather than \texttt{TIMESTAMP} columns.
            This avoids timezone conversion issues, as epoch values are inherently UTC,
            and simplifies interoperability with the \texttt{kotlinx.datetime.Instant}
            type used throughout the backend.

      \item \textbf{Serial primary keys}: All tables use \texttt{SERIAL}
            (auto-incrementing integer) primary keys rather than UUIDs. This simplifies
            joins and keeps index sizes small, which is well-suited to the expected data
            volume of the platform.
\end{itemize}
