%
% Chapter — Deployment
%
\chapter{Deployment} \label{cap:deployment}

This chapter describes how the components introduced in the previous chapters are
packaged, provisioned, and run in production. Whereas Chapter~\ref{cap:architecture}
presented the \emph{logical} architecture --- how the source code is organised into
modules and layers --- this chapter presents the \emph{physical} deployment view: the
container images, the services that run on the host, the reverse-proxy layers, and the
path an external request travels to reach the application. The two views are
complementary: the Spring Boot backend that appears as a single logical component in
Chapter~\ref{cap:architecture} is, at deployment time, one container image among
several orchestrated together.

The platform runs on a single virtual machine (VM). Almost every runtime component
runs as a Docker~\cite{docker} container, orchestrated through Docker Compose; the only exception is
the outermost reverse proxy, a system-level Nginx~\cite{nginx} provisioned directly on the host to
terminate HTTPS. The complete topology is shown in Figure~\ref{fig:deployment}
(Appendix~\ref{app:deployment}).

\section{Container Images} \label{sec:dep_images}

The system is packaged as three application images, complemented by the official
SeaweedFS image used for the object-storage cluster:

\begin{description}
      \item[\texttt{diariolx-jvm}] Built from \texttt{eclipse-temurin:21}. The Spring
            Boot application is split into three Docker layers --- external libraries,
            \texttt{META-INF} metadata, and the application classes --- so that the
            least-frequently-changed files occupy the lower, cacheable layers and only
            the application layer is rebuilt on most changes. The entry point launches
            \texttt{pt.ipl.diariolx.DiarioLXApplicationKt}.

      \item[\texttt{diariolx-web}] A multi-stage image that bundles \emph{both} the
            frontend and the reverse proxy. The first stage (\texttt{node:22-alpine})
            installs the frontend dependencies with \texttt{npm ci} and produces the
            React production build (\texttt{npm run build}) in \texttt{/app/dist}. The
            second stage (\texttt{nginx:alpine}) copies the compiled build into the web
            root (\texttt{/usr/share/nginx/html}) and installs the shared Nginx
            configuration. Because
            the build context is the repository root, a single image has access to both
            the frontend sources and the shared Nginx configuration.

      \item[\texttt{diariolx-postgres}] A PostgreSQL image pre-provisioned with the
            database schema --- tables, indexes, and views --- described in
            Chapter~\ref{cap:datamodel}.

      \item[SeaweedFS] The official \texttt{chrislusf/seaweedfs} image, used unchanged
            for the object-storage cluster.
\end{description}

\section{Orchestrated Services} \label{sec:dep_services}

Docker Compose orchestrates the following services on the VM:

\begin{itemize}
      \item \textbf{PostgreSQL} --- the relational database; its data is kept on a named
            volume so it survives container recreation, and its configuration is read
            from \texttt{.env}.

      \item \textbf{SeaweedFS cluster} --- the self-hosted, S3-compatible object storage
            for multimedia assets, split into cooperating roles: a \emph{master}
            (cluster coordination), a \emph{volume} server (stores the file bytes), a
            \emph{filer} (filesystem semantics over the volumes), an \emph{S3 gateway}
            (exposes the S3 API on port 8333, configured through
            \texttt{seaweed-config.json}), and an \emph{admin} service. The master,
            volume, and filer each persist to their own named volume.

      \item \textbf{Spring Boot backend} (\texttt{diariolx-jvm-dynamic}) --- the
            application server, with its environment supplied by \texttt{.env}.

      \item \textbf{Web} (\texttt{diariolx-web}) --- the combined frontend-and-proxy
            image, published to the host and declared to depend on the S3 gateway and the
            backend so that they start first.
\end{itemize}

Persistence is achieved entirely through named Docker volumes (one for PostgreSQL and
one each for the SeaweedFS master, volume, and filer), which decouple the stored data
from the lifecycle of the containers.

\section{Reverse Proxy and Routing} \label{sec:dep_proxy}

Requests pass through two Nginx layers, each with a distinct responsibility.

The \textbf{host Nginx} is a system service provisioned directly on the VM and managed
by Certbot. It listens on port 443 using a Let's Encrypt~\cite{letsencrypt} certificate issued for
\texttt{diariolx.duckdns.org}, redirects all plaintext traffic on port 80 to HTTPS,
and forwards the decrypted requests to the containerised Nginx on
\texttt{localhost:8180}, preserving the original host and client information through the
\texttt{X-Forwarded-*} headers. Dynamic DNS is provided by DuckDNS, which maps
\texttt{diariolx.duckdns.org} to the VM's public IP address. This layer is not part of
the Compose file, since TLS termination and certificate renewal are host-level
concerns.

The \textbf{containerised Nginx} --- the \texttt{diariolx-web} image, listening on port
8180 --- is the single entry point into the Compose network and performs path-based
routing:

\begin{itemize}
      \item \texttt{/} --- serves the React static build, falling back to
            \texttt{index.html} (\texttt{try\_files}) so that client-side routing works
            on deep links;
      \item \texttt{/api} and \texttt{/docs} --- proxied to the Spring Boot backend (the
            REST API and the Scalar API documentation, respectively);
      \item \texttt{/media/} --- proxied to the SeaweedFS S3 gateway, forwarding the
            host and client headers required for object access.
\end{itemize}

As a special case of the first rule, requests from social-media crawlers are rewritten to
a server-rendered preview endpoint rather than served the SPA shell; this is covered in
Section~\ref{sec:dep_previews}.

Beyond routing, this Nginx layer applies a tiered caching policy that offloads both the
backend and the object store. Responses from the public API are held in a short
cache of roughly thirty seconds, which still absorbs load on hot endpoints such as the
homepage while keeping the window in which content can appear stale small enough to be
imperceptible to readers. Media objects, whose paths are
content-addressed and never change, are cached for thirty days and marked
\texttt{immutable}, as are the static frontend assets, which carry content-hashed file
names. The backoffice, authentication, and documentation endpoints are explicitly marked
\texttt{no-store}, so that no authenticated or administrative response is ever cached.
Each response exposes an \texttt{X-Cache-Status} header indicating whether it was served
from cache.

\section{Social Media Previews} \label{sec:dep_previews}

For a digital newspaper, the sharing of articles on social media is an important channel
for reaching readers, and an inviting link preview --- a title, a summary, and an image ---
is what turns a shared link into a visit. Delivering good previews, however, is at odds
with a client-rendered single-page application: because the public site renders in the
browser, its HTML shell carries no per-article metadata, so a crawler fetching an article
URL would find nothing from which to build a preview.

To resolve this, the containerised Nginx inspects the \texttt{User-Agent} of requests to
article paths (\texttt{/p/\{slug\}}) and, when it matches a known social-media crawler ---
such as those of Facebook, WhatsApp, X, or LinkedIn --- rewrites the request to a dedicated
backend endpoint that returns a lightweight server-rendered page carrying the Open
Graph~\cite{ogp} metadata (title, description, image, and canonical URL) for that article.
Human visitors are unaffected and continue to receive the single-page application.

This is a targeted solution rather than full server-side rendering: it covers only the
crawler user agents that are explicitly recognised, and both widening that set and, more
fundamentally, moving to genuine server-side rendering of the public pages remain possible
improvements.

\section{Request Lifecycle} \label{sec:dep_lifecycle}

A typical request therefore flows as follows. The browser resolves
\texttt{diariolx.duckdns.org} through DuckDNS to the VM and issues an HTTPS request. The
host Nginx terminates TLS and forwards the request to the containerised Nginx, which
either serves the static frontend directly, proxies \texttt{/api} (and \texttt{/docs})
to the backend, or proxies \texttt{/media/*} to the SeaweedFS S3 gateway. The backend,
in turn, reads from and writes to PostgreSQL and issues object-storage operations
against SeaweedFS. This single-entry-point design keeps the internal services
unreachable from outside the VM: only the host Nginx is exposed to the public network.

\section{Continuous Delivery} \label{sec:dep_cd}

The application images are built and published to the GitHub Container Registry (GHCR)~\cite{ghcr}
by a GitHub Actions~\cite{github_actions} workflow, so that the server never builds anything itself. On each
release, the workflow builds the three application images (\texttt{diariolx-jvm},
\texttt{diariolx-web}, and \texttt{diariolx-postgres}) and pushes them to GHCR. Docker
layer caching keeps rebuilds fast by reusing unchanged layers across runs --- most
notably the frontend dependency and build layers of the \texttt{diariolx-web} image.

\section{Running in Production} \label{sec:dep_run}

Deploying on the VM reduces to pulling the published images and starting the stack; the
server itself builds nothing (Section~\ref{sec:dep_cd}). The production Compose
file references the images by tag and reads its configuration from a \texttt{.env} file.
The secrets --- that \texttt{.env} file and the \texttt{seaweed-config.json} that
pre-declares the S3 identities (so no manual object-storage configuration is needed) ---
are provisioned directly on the VM and are never baked into the published images. The
deployment is deliberately minimal in its exposed surface: only the entry-point
\texttt{diariolx-web} container is published to the host, so the internal services
(PostgreSQL, the SeaweedFS S3 gateway,
and the admin service) are reachable exclusively over the Compose network. Because the
images are self-contained and all data lives on named volumes, the stack can be
recreated without loss of stored content.
