%
% Chapter — Conclusion
%
\chapter{Conclusion} \label{cap:conclusion}

This report has presented the design and implementation of \textit{Diário LX}, a
purpose-built digital journalism platform developed in collaboration with the Escola
Superior de Comunicação Social (ESCS) and the LIACOM research unit. The project set out
to support the complete editorial lifecycle --- from content authoring to publication
--- through a role-based workflow, a structured multimedia content model, and a
dedicated public website, within a decoupled client--server architecture. As discussed
in Chapter~\ref{cap:background}, existing content management systems such as WordPress
and Ghost were evaluated but did not fully accommodate these requirements, motivating
the development of a tailored solution. This document describes the final version of
that system, which has been implemented and deployed.

\section{Summary of Work} \label{sec:conc_summary}

The delivered platform realises the architecture and requirements described in the
preceding chapters:

\begin{itemize}
      \item \textbf{Editorial workflow}: the full content lifecycle --- draft, pending
            review, approved, and rejected --- is implemented and enforced according
            to the actor's role. Contributors submit content for review; Editors and
            Administrators approve it (publishing it) or reject it with a
            comment, and every transition is recorded in the content's history.

      \item \textbf{Content model and editor}: the backoffice content editor supports
            block-based authoring across all content types --- articles, videos,
            podcasts, episodes, and photo essays --- with inline formatting, single media,
            captioned image galleries, external embeds, and full metadata configuration.
            Media assets carry credits and accessible descriptions.

      \item \textbf{Backoffice}: the editorial interface covers content management, user
            administration, invitation-based onboarding, category and tag management, the
            media library, and curation of the public homepage.

      \item \textbf{Public website}: the reader-facing site presents a homepage driven by
            the backoffice curation, article pages, category, tag, and section listings,
            dedicated podcast, episode, and photo-essay pages, and team and author pages,
            with a responsive layout.

      \item \textbf{Security}: authentication uses short-lived JWT access tokens and
            long-lived opaque refresh tokens, issued as \texttt{HttpOnly},
            \texttt{SameSite=Strict}, and --- in production --- \texttt{Secure} cookies.
            Role-based access control is enforced at the HTTP boundary.

      \item \textbf{Deployment}: the system is fully containerised and its images are
            published to a container registry through continuous integration, allowing
            the production server to deploy by pulling rather than building, behind a
            TLS-terminating reverse proxy (Chapter~\ref{cap:deployment}).

      \item \textbf{Quality assurance}: automated backend and end-to-end frontend tests
            run in a continuous integration pipeline, and the REST API is documented
            through an interactive OpenAPI interface.
\end{itemize}

\section{Future Work} \label{sec:conc_future}

\textit{Diário LX} was conceived with a twofold purpose: to deliver a functional product
that the ESCS editorial team can adopt and use in practice, and to provide a foundation
that other developers can maintain and extend once this project concludes. This second
goal deliberately shaped the engineering approach --- the strict separation into
independent modules and layers, the explicit and readable data-access code, the
documented REST API, the automated test suite, and the fully containerised,
registry-based deployment. Together, these lower the barrier for a new contributor to
understand the system, run it locally, and build on it with confidence, so that the work
delivered here can be continued rather than restarted.

On this basis, several enhancements are envisaged as natural continuations of the work:

\begin{itemize}
      \item \textbf{Internationalisation and content translation}: the frontend already
            centralises the translation of the interface's static text; the natural next
            step is to extend multi-language support to the editorial content itself,
            allowing a publication to be authored in more than one language and served to
            each reader in their preferred one.

      \item \textbf{Cache invalidation}: complementing the current time-based caching ---
            where public responses are held only briefly to bound how long content can
            appear stale --- with explicit invalidation, so that approving or editing a
            publication purges the affected entries and the change is reflected immediately,
            without having to keep the cache window short.

      \item \textbf{Search engine optimisation}: building on the Open Graph previews
            already served to social-media crawlers (Section~\ref{sec:dep_proxy}) by
            enriching public pages with structured data and a generated sitemap, and,
            more broadly, adopting server-side rendering so that all content is fully
            indexable rather than only prerendered for recognised crawlers.

      \item \textbf{Asynchronous media processing}: offloading heavy media handling to
            background tasks so that uploads return promptly instead of blocking on large
            files --- generating server-side resized, responsive variants of images and
            extracting thumbnail frames for videos, rather than serving a single original of
            each asset.

      \item \textbf{Accessibility}: further accessibility improvements, such as audio
            description for multimedia content, broadening the platform's reach.

      \item \textbf{Content recommendations}: surfacing related articles on the article
            page to improve navigation and reader retention.
\end{itemize}

Because the platform is delivered as a working, deployed system built on a modular and
well-documented codebase, each of these extensions can be approached incrementally,
without a structural rewrite --- which was, from the outset, an explicit objective of the
project.

\section{Final Remarks} \label{sec:conc_remarks}

\textit{Diário LX} demonstrates the feasibility of a purpose-built solution tailored to a
specific editorial context, where a general-purpose content management system would have
imposed structural and workflow compromises. The resulting platform delivers a complete,
deployed system that covers the editorial process end to end, together with the public
experience through which its content is read. Beyond the product itself, the project
provided practical experience in full-stack development, layered software architecture,
functional error handling, and containerised deployment, applied to the concrete
requirements of a real editorial partner.
